---
title: "Capofari_Week_3_IS607"
author: "Nicholas Capofari"
date: "September 11, 2015"
output: html_document
---

This week's assignment is to transform a data set for easier downstream analysis. I understand that the assignment required the use of a data set from the UCI Machine Learning Repository, but I felt that since the football season is upon us the following data set would be pertinent.  I hope this is ok.

This data set can be found at:  https://raw.githubusercontent.com/fivethirtyeight/data/master/nfl-suspensions/nfl-suspensions-data.csv.  
http://fivethirtyeight.com is an interesting site for the data enthusiast.  

My original .Rmd file can be found at http://google.com.

```{r, include=FALSE}
library(RCurl)
x <- getURL("https://raw.githubusercontent.com/fivethirtyeight/data/master/nfl-suspensions/nfl-suspensions-data.csv")
y <- read.csv(text = x)
```
The data set I chose lists all NFL suspensions from ```r min(y$year, na.rm = TRUE)``` to ```r max(y$year, na.rm = TRUE)```.  Here are the first 5 observations from the data set: 

```{r, echo=FALSE}
head(y)
```

For my purposes I want to focus on all substance abuse related suspensions from 1980 to ```r max(y$year, na.rm = TRUE)```.  All superflorous columns will be removed.

```{r, echo=FALSE}
trans_data <- subset(y, y$year > 1979 & y$category == "Substance abuse")
keeps <- c("name", "team", "year", "games")
trans_data <- trans_data[keeps]
trans_data <- transform(trans_data, as.numeric(trans_data$Length))
names(trans_data) <- c("Name", "Team", "Year", "Length")
```

Here are the first 5 observations from the transformed data set:
```{r, echo=FALSE}
head(trans_data, addrownums = FALSE)
```
And now it is much easier to work with the data:
```{r, echo=FALSE}
yrs <- max(trans_data$Year) - min(trans_data$Year)
hist(trans_data$Year, main = "NFL Suspensions", sub = "1980-2014", xlab = "Years", ylab = "Suspensions", breaks = yrs)
```


