---
title: "Capofari_Week_4_IS607"
author: "Nicholas Capofari"
date: "September 18, 2015"
output: html_document
---
```{r, echo=FALSE}
library(stringr)
```
###Example 4
a) The regular expression [0-9]+\\$ takes the first instance of 1 or more digits concatenated with 1 $ symbol.
```{r}
unlist(str_extract_all("Bobby won 314.15$ and Alice won 271$.", "[0-9]+\\$"))
```

b) The regular expression \\b[a-z]{1,4}\\b matches strings of length 1, 2, 3, or 4, that are made of *only* lower case letters.
```{r}
unlist(str_extract_all("In my younger and more vulnerable years my father gave me some advice that I've been turning over in my mind ever since.", "\\b[a-z]{1,4}\\b"))
```

c) The regular expression .*?\\.txt$ matches all possible strings that end with .txt.
```{r}
sample_text <- c("my_text_file.txt", "my_python_file.py", "my other text file.txt")
unlist(str_extract(sample_text, ".*?\\.txt$"))
```

d) The regular expression \\d{2}/\\d{2}/\\d{4} matches dates in the form ##/##/####.
```{r}
sample_text <- c("03/14/2015", "1/10/2011", "08/25/81")
unlist(str_extract(sample_text, "\\d{2}/\\d{2}/\\d{4}"))
```
###Example 5
[0-9]+\\$ can be rewritten as\\d(\\d)*\\${1}.  These two regular expressions will perform the same task.
```{r}
unlist(str_extract_all("Bobby won 271$ and Alice won 314.15$ and Chris won 1$.", "\\d(\\d)*\\${1}"))
```
###Example 6
chunkylover[at]aol[dot]com
```{r}
e_mail <- "chunkylover53[at]aol[dot]com"
e_mail <- str_replace(str_replace(e_mail, "\\[dot\\]", "."), "\\[at\\]", "@")
e_mail
unlist(str_extract_all(e_mail, "[:digit:]"))
```
The expression \\D matches all characters are *not* digits.  we should use \\d.
```{r}
unlist(str_extract_all(e_mail, "\\D"))
unlist(str_extract_all(e_mail, "\\d"))
```


